Model validation plots

### 2021-08-11 doc/validate_models.Rmd
re-organize validation

---
title: "Model validation plots"
author: "Eric R. Scott"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

@@

# TODO

@@

# Analysis

## Survival

Quantile residuals are more appropriate for binomial models.

### Fragments

```{r}
withr::with_dir(here::here(),{
  tar_load(c(s_1ha, s_cf, s_cf_sub))
})
```

@@

```{r echo=TRUE}
s_1ha_res <- qresid(s_1ha)
par(mfrow = c(2,2))
#histogram
plot(density(s_1ha_res))
#QQ plot
qqnorm(s_1ha_res); qqline(s_1ha_res)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(s_1ha), residuals(s_1ha, type = "response"))
```
```{r}
# gam.check(s_1ha)
```

Basis dimension checking with `gam.check()` doesn't appear to work for `dlnm` crossbasis smooths. Instead I'll use a method from `?choose.k` to check for adequate knots.  Unfortunately, `bs = "cs"` doesn't work with `dlnm`, so I'll use `select = TRUE` instead to reduce chance of overfitting.

### Continuous Forest

@@

```{r echo=TRUE}
s_cf_res <- qresid(s_cf)
par(mfrow = c(2,2))
#histogram
plot(density(s_cf_res))
#QQ plot
qqnorm(s_cf_res); qqline(s_cf_res)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(s_cf), residuals(s_cf, type = "response"))
```

### Continuous forest subsample

To check that differences are not purely due to sample size differences, particularly that lower edf in continuous forests is due to higher sample size.  For survival, edf is actually higher in CF with a larger sample size.  With subsample, edf are more similar.

@@

```{r}
rm(s_1ha, s_cf, s_cf_sub)
```

## Growth

```{r}
withr::with_dir(here::here(),{
  tar_load(c(g_1ha, g_cf, g_cf_sub))
})
```

### Fragments

@@

```{r}
appraise(g_1ha)
```
```{r}
g_1ha_res <- residuals(g_1ha)
# looking for near zero edf
gam(g_1ha_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(g_1ha)) #fine

### Continuous Forest

@@

### Continuous Forest subsample

@@

```{r}
rm(g_1ha, g_cf, g_cf_sub)
```

## Flowering

Quantile residuals are more appropriate for binomial models.

```{r}
withr::with_dir(here::here(),{
  tar_load(c(f_1ha, f_cf, f_cf_sub))
})
```

### Fragments

@@

### Continuous Forest

@@

### Flowering prob of largest plants

Figure out flowering probability of largest quartile of plants in each habitat

```{r}
newdf_cf <-
  f_cf$model %>% #use model data frame
  mutate(quartiles = ntile(log_size_prev, 4)) %>% 
  filter(quartiles == 4)

newdf_1ha <-
  f_1ha$model %>% 
  mutate(quartiles = ntile(log_size_prev, 4)) %>% 
  filter(quartiles == 4)
```

```{r}
#equivalent to flowering once every __ years
1/predict(f_cf, newdata = newdf_cf) %>% mean() %>% plogis()
1/predict(f_1ha, newdata = newdf_1ha) %>% mean() %>% plogis()
```

```{r}
rm(f_1ha, f_cf, f_cf_sub)
```

>>>>

---
title: "Model validation plots"
author: "Eric R. Scott"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

@@

**TODO**

@@

# Survival

```{r, echo=FALSE}
withr::with_dir(here::here(),{
  tar_load(c(s_1ha, s_cf, s_cf_sub))
})
```

## Forest Fragments

### Diagnostics

```{r echo=TRUE}
s_1ha_res <- qresid(s_1ha)
par(mfrow = c(2,2))
#histogram
plot(density(s_1ha_res))
#QQ plot
qqnorm(s_1ha_res); qqline(s_1ha_res)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(s_1ha), residuals(s_1ha, type = "response"))
```

### Basis dimensions

```{r}
# gam.check(s_1ha)
```

Basis dimension checking with `gam.check()` doesn't appear to work for `dlnm` crossbasis smooths. Instead I'll use a method described in the help file `?choose.k` to check for adequate knots.  Unfortunately, `bs = "cs"` doesn't work with `dlnm`, so I'll use `select = TRUE` instead to reduce chance of overfitting.

### Summary

@@

## Continuous Forest

### Diagnostics

```{r echo=TRUE}
s_cf_res <- qresid(s_cf)
par(mfrow = c(2,2))
#histogram
plot(density(s_cf_res))
#QQ plot
qqnorm(s_cf_res); qqline(s_cf_res)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(s_cf), residuals(s_cf, type = "response"))
```
### Basis dimensions

### Summary

@@

## Effect of sample size

Here I use a random sub-sample to check that differences between continuous forest and fragments are not purely due to sample size differences, particularly differences in the complexity of the crossbasis smooth.

@@

```{r echo=FALSE}
rm(s_1ha, s_cf, s_cf_sub)
```

# Growth

```{r echo=FALSE}
withr::with_dir(here::here(),{
  tar_load(c(g_1ha, g_cf, g_cf_sub))
})
```

## Fragments

### Diagnostics

```{r}
appraise(g_1ha)
```

### Basis dimensions

```{r}
g_1ha_res <- residuals(g_1ha)
# looking for near zero edf
gam(g_1ha_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(g_1ha)) #fine

### Summary

@@

## Continuous Forest

### Diagnostics

### Basis dimensions

### Summary

@@

## Continuous Forest subsample

@@

```{r echo=FALSE}
rm(g_1ha, g_cf, g_cf_sub)
```

# Flowering

```{r echo=FALSE}
withr::with_dir(here::here(),{
  tar_load(c(f_1ha, f_cf, f_cf_sub))
})
```

## Fragments

### Diagnostics

### Basis dimensions

### Summary

@@

## Continuous Forest

### Diagnostics

### Basis dimensions

### Summary

@@
